Breareal

### 2022-08-08 02-overview.Rmd
proj6
WW

# Overview {-}

<!-- This page is useful because by calling the indicator rmd files as child files we bypass the need to sequentially numbering them. An overview page is anywhays useful.  -->

## Areal av isbreer / Glacier area {-}
This indicator is about ....

@@

```{r, child= c('indicators/indicatorTest.Rmd')}
```

```{r, child= c('indicators/indicatorTest2.Rmd')}
```

>>>>

# Overview of indicators {-}

## Areal av isbreer / Glacier area {-}
This indicator reflects the relative loss of glacial area between tha last climatic normal persion and today.

@@

### 2022-08-08 breareal.Rmd
WW

# (PART\*) Indicators from the alpine assessment {-}

# Areal av isbreer {-}

```{r, include=FALSE, message=FALSE}
library(knitr)
library(sf)
library(terra)
library(raster) # only needed for creating boundary boxes, which can be done using sf
library(dplyr)
library(lwgeom)
library(tmap)
library(dplyr)
#library(ggplot2)
#library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Introduction
This indicator describes the relative change in glacial extent for each of fire regions in Norway.

@@

## Ecosystem characteristic
The indicator was initially assigned the _Abiotiske forhold_ characteristic, but now it is considered under _Landskapsøkologiske mønstre_. The initial justificaton was that glacial melting would influence water availability downstream. Now we put more weight on glaciers as not just resevoirs of ice and water, but as a habitat, and the presence or abundance of a specific habitat can be though of as an aspect of landscape state.

The indicator is similarly assiged to the SEEA ECT class _C1 Landscape characteristics_.

@@

## References:
* Andreassen, L.M., & Winsvold, S.H. (eds.). 2012. Inventory of Norwegian glaciers. NVE Rapport 38, Norges Vassdrags- og energidirektorat, 236 s

@@

#### Glacial atlas
This is the new glacial atlas from 2018 and 2019 based on Sentinell (ref. Liss Marie Andreassen, NVE)
```{r}
# This first part is if you are running this from the NINA R server. 
#breatlas <- sf::st_read('/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/breatlas_2018_2019/Breatlas_20182019_temp20210922_lma.shp')

breatlas <- sf::st_read('P:/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/breatlas_2018_2019/Breatlas_20182019_temp20210922_lma.shp')
```

@@

```{r}
#n50 <- sf::st_read('/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/n50/cryoclim_GAO_NO_1952_1985_UTM_33N.shp')%>%
#  st_transform(crs = crs(breatlas))

n50 <- sf::st_read('P:/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/n50/cryoclim_GAO_NO_1952_1985_UTM_33N.shp')%>%
  st_transform(crs = crs(breatlas))

@@

Plotting an interactive map with scaled indicator values
```{r}
tmap_mode("view")
tm_shape(reg_clipped) + 
  tm_polygons(col="skalert_indikator", 
              border.col = "white")+
  tm_text("skalert_indikator_r", size=1.5)+
  tm_shape(nor)+
  tm_polygons(alpha = 0,border.col = "black")
```

@@

```{r, eval=F}
sf::st_write(exp, "indicatorMaps/breareal.shp")
```

>>>>

# (PART\*) INDICATORS FROM THE ALPINE ASSESSMENT {-}

# Areal av isbreer

```{r, include=FALSE, message=FALSE}
library(knitr)
library(sf)
library(terra)
library(raster) # only needed for creating boundary boxes, which can be done using sf
library(dplyr)
library(lwgeom)
library(tmap)
library(dplyr)
#library(ggplot2)
#library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```
<br />

_Author and date:_
Anders Kolstad

```{r, echo=F}
Sys.Date()
```

@@

## Introduction
This indicator describes the relative change in glacial extent for each of fire regions in Norway. A [slightly different indicator](https://ninanor.github.io/IBECA/faktaark.html#Areal_av_isbreer) was used in the alpine assessment.

@@

## Ecosystem characteristic
### Norwegian standard
The indicator was initially assigned the _Abiotiske forhold_ characteristic, but now it is considered under _Landskapsøkologiske mønstre_. The initial justificaton was that glacial melting would influence water availability downstream. Now we put more weight on glaciers as not just resevoirs of ice and water, but as a habitat, and the presence or abundance of a specific habitat can be though of as an aspect of landscape state.

### SEEA ECT
The indicator is similarly assiged to the SEEA ECT class _C1 Landscape characteristics_.

@@

## References
* Andreassen, L.M., & Winsvold, S.H. (eds.). 2012. Inventory of Norwegian glaciers. NVE Rapport 38, Norges Vassdrags- og energidirektorat, 236 s

@@

#### Glacial atlas
This is the new glacial atlas from 2018 and 2019 based on Sentinell (ref. Liss Marie Andreassen, NVE)
```{r}

# server path
breatlas_path <- '/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/breatlas_2018_2019/Breatlas_20182019_temp20210922_lma.shp'

# local path 
#breatlas_path <- 'P:/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/breatlas_2018_2019/Breatlas_20182019_temp20210922_lma.shp'

breatlas <- sf::st_read(breatlas_path)
```

@@

# server path
n50_path <- '/data/P-Prosjekter/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/n50/cryoclim_GAO_NO_1952_1985_UTM_33N.shp'

# local path
#n50_path <- 'P:/41201042_okologisk_tilstand_fastlandsnorge_2020_dataanaly/fjell2021/data/breatlas/n50/cryoclim_GAO_NO_1952_1985_UTM_33N.shp'

n50 <- sf::st_read(n50_path)%>%
  st_transform(crs = crs(breatlas))
```

@@

Plotting a static map with scaled indicator values
```{r}
tmap_mode("plot") # mode 'view' takes a lot longer to render, but looks better
tm_shape(reg_clipped) + 
  tm_polygons(col="skalert_indikator", 
              border.col = "white")+
  tm_text("skalert_indikator_r", size=1.5)+
  tm_shape(nor)+
  tm_polygons(alpha = 0,border.col = "black")
```

@@

```{r, eval=F}
sf::st_write(exp, "indicators/indicatorMaps/breareal.shp")
```
